The resurgence of the linear optics interferometer --- recent advances \& applications

### 2017-05-11 paper.tex
Cleaned up Section 6.3.

Broadly speaking, there are two main classes of photodetectors: non-photon-number-resolving (also referred to as bucket, or on/off) detectors, and photon-number-resolving (PNR) detectors. As the names suggest, non-PNR detectors only detect the presence or absence of photons in an incident beam, while PNR detectors are able to measure its number of photons. The outcomes of a PNR detector are described by the measurement projectors,
\begin{align}
\hat\Pi_n &= \ket{n}\bra{n}, \ n=0,1,\ldots \ ,
\end{align}
while those of the non-photon-number-resolving detectors are given by
\begin{align}
\hat\Pi_\text{off} &= \hat\Pi_0, \nonumber \\
\hat\Pi_\text{on} &= \hat{I} - \hat\Pi_\text{on}\ .
\end{align}
The dominant challenge facing the experimental implementation of both types of detectors is inefficiency due to loss, which results in higher photon-number being confused as having lower photon-number. In some protocols, such as type-II fusion, which relies on coincidence detection, inefficiency is not problematic. However, in most applications, including type-I fusion, this is highly problematic as photon-number confusion results in decoherence via projecting onto the wrong photon-number, unknown to the observer. Other common problems that plague these detectors are dark counts which add spurious counts from the background, and dead-time which causes the detector to lose counts during a period when it is inactive immediately after the detection of photons.

\subsubsection{Non-photon-number-resolving (non-PNR) detectors}
The first type of non-PNR detectors are photomultiplier tubes (PMTs). A PMT comprises of a vacuum tube with a photocathode for light absorption. The photoelectric effect then releases electrons in the photocathode, and this current is multiplied by a cascade of secondary electron emissions from a series of electrodes. This amplified signal can be picked up to indicate the presence of photons.

In applications for quantum information processing, a well-established alternative to PMTs is the avalanche photodiode (APD). These are devices made up of semiconductor material, usually silicon, which is structured as a p-n junction and put into Geiger mode operation. When a photon is incident on the APD, electrons that are generated undergo an avalanche and the current that is generated can be measured. APDs are affordable, offer low dark count rates with reasonable efficiency, and can be made to operate in wavelengths in the visible to near-infrared range. A downside to using APD is their deadtime. This is a period of time following an avalanche during which the avalanche is stopped by a quenching circuit to reset the APD. No photons can be detected during the deadtime and contributes to photon loss in an APD.

The last type of non-PNR detectors we discuss is the superconducting nanowire single-photon detector (SNSPD) which have recently become commercially available. This detector uses a narrow superconducting wire that is carrying a current operating at just below the critical current density. Above this current density, the superconducting wire would undergo a phase change back to its normal resistance which happens when it absorbs an incident photon. The current in the wire would then by-pass it to go through the neighboring region which is still superconducting, and therefore at a lower resistance. This causes a voltage spike in the neighboring region which can be detected.

\subsubsection{Photon-number-resolving detectors}

Multiple technologies for realizing PNR detectors already exist and have been described by multiple reviews \cite{bib:Hadfield09,bib:Eisaman11}, so we shall only outline a few notable examples here. One of these is the superconducting transition edge sensor (TES) . In TES, a layer of superconducting material is operated at a temperature in its transition between superconducting and normal resistance. Any small change in temperature will thus cause a large change in resistance of this layer which can be measured. This happens when photons are absorbed, and heats up the superconductor. Such detectors can be made so sensitive that the energy of single photons can be detected. Another advantage of the TES is its high efficiency which can be made very close to 100\% \cite{bib:Rosenberg05}, and low dark count rates. However, they are quite slow, have low maximum counting rates, and are costly to operate due to their very low operating temperatures.

The visible light photon counter (VLPC) is another kind of PNR detector. The way VLPC works is in principle very similar to the APD. The top layer where the photon impinges is made of silicon, followed by another gain layer that is lightly doped with arsenic (As). When a photon is incident on this detector, it creates electrons which are then accelerated into the gain layer and ionize the As impurities, thereby exciting and scattering more electrons. An avalanche multiplication ensues just like in an APD. However, owing to the presence of the As gain later, the VLPC has a low multiplication noise which results in a single photon absorption event creating an electrical signal that is always of the same magnitude. Thus, the output electrical signal is just proportional to the number of detected photons.

Last, we discuss an approach to creating PNR detectors from non-PNR detectors by connecting a bunch of the latter in parallel to one another, and summing the signals at the output to give a combined signal that is proportional to that of just one. Such an approach has been demonstrated with SNSPD connected in parallel \cite{bib:Divochiy08}, and APDs connected in an array that can be either on the same semiconductor substrate, or as an interferometric architecture \cite{bib:Afek09,bib:Chrapkiewicz14,bib:Chen17}. A drawback of the parallel-SNSPDs is their low quantum efficiency, while the spatially multiplexed APDs suffer from spurious counts known as crosstalk that arises from avalanches triggered by electrons launched from neighboring APDs \cite{bib:Eraerds07,bib:Afek09,bib:Kalashnikov12}.

@@

\subsection{Linear optics quantum computation (LOQC)}
With the advancement in single-photon state preparation, detectors and quantum interferometers, quantum computation using linear optics has become feasible. 
For universal quantum computation using linear optics and single photon inputs, nonlinear couplings between the different optical modes are needed. Photons, being bosons, do not interact with one another like the way fermions do. Hence, some careful engineering has to be done in order to achieve this nonlinear coupling between modes. Some of the methods used to introduce this nonlinearity include measurements and feedforwarding \cite{bib:KLM01}, and  photon-atom interactions \cite{bib:Brod16}. These nonlinearities have been used to construct entangling gates required for (approximate) universality in quantum computation using the circuit model. They have also been used to construct entangled graph states, like the cluster states covered in Section \ref{section:cluster}, that are used in MBQC. Comprehensive reviews have been written previously to cover this topic \cite{bib:Kok05,bib:JW12}, hence we will skip some basic information to focus on recent advances. There are multiple challenges to implementing a scalable, fully-optical quantum computer, and progress has been focused on overcoming these challenges. We will discuss a few of these challenges, and the advances that have been made to overcome them.

@@

 \begin{figure}[!htb]
 \centering
\includegraphics[width=0.4\columnwidth]{phE}
\caption{A schematic diagram of the Mach-Zehnder interferometer comprising of two 50/50 beamsplitters with a phase shift $\varphi$ in one of its two paths.} \label{fig:MZ}	
\end{figure}

>>>>

Broadly speaking, there are two main classes of photodetectors: non-photon-number-resolving (also referred to as bucket, or on/off) detectors, and photon-number-resolving (PNR) detectors. As the names suggest, non-PNR detectors only detect the presence or absence of photons in an incident beam, while PNR detectors are able to measure its number of photons. The outcomes of a PNR detector are described by the measurement projectors,
\begin{align}
\hat\Pi_n &= \ket{n}\bra{n}, \ n=0,1,\ldots \ ,
\end{align}
while those of the non-photon-number-resolving detectors are given by
\begin{align}
\hat\Pi_\text{off} &= \hat\Pi_0, \nonumber \\
\hat\Pi_\text{on} &= \hat{I} - \hat\Pi_\text{on}\ .
\end{align}
Detailed reviews have been written about these detectors \cite{bib:Hadfield09,bib:Eisaman11}, so we shall only outline a few notable examples here.

A common example of a non-PNR detector is the avalanche photodiode (APD). These devices are made up of semiconductor material, usually silicon, which generate an electron that quickly generates a cascade of electrons. This avalanche of electrons generate a current that can be measured. APDs are affordable, offer low dark count rates with reasonable efficiency, and can be made to operate in a broad spectrum of wavelengths; from the visible to near-infrared range
% A downside to using APD is their deadtime. This is a period of time 
%following an avalanche during which the avalanche is stopped by a 
%quenching circuit to reset the APD. No photons can be detected during the 
%deadtime and contributes to photon loss in an APD. 
Superconducting nanowire single-photon detector (SNSPD) is another  type of non-PNR detector that has recently become commercially available. This detector uses a narrow superconducting wire that carries a current just below the critical current density, above which the superconducting wire would undergo a phase change back to its normal resistance. This happens when the SNSPD absorbs an incident photon, and the current in the wire would then by-pass it to go through the neighboring region which is at a lower resistance. The voltage spike that ensues is measured to give a photon number count.

PNR detectors are a wonderful addition to the toolbox of linear optics. A notable example is the superconducting transition edge sensor (TES) which has an extremely high quantum efficiency. The principle it operates on is very similar to the SNSPD, but with a much higher sensitivity that allows it to detect the energy of single photons.
%low dark count rates. However, they are quite slow, have low maximum 
%counting rates, and are costly to operate due to their very low operating 
%temperatures. 
The visible light photon counter (VLPC) is another kind of PNR detector. The way VLPC works is in principle very similar to the APD, but with an additional layer of silicon that is lightly doped with arsenic (As). Owing to the presence of the As, a single photon absorption event always create an electrical signal that is always of the same magnitude. Thus, the output electrical signal is just proportional to the number of detected photons.
Last, PNR detectors can be created by connecting many non-PNR detectors in parallel to one another, and summing the signals at the output. Such an approach has been successfully demonstrated with SNSPD \cite{bib:Divochiy08}, and APDs \cite{bib:Afek09,bib:Chrapkiewicz14,bib:Chen17}.

The dominant challenge facing the experimental implementation of both types of detectors is inefficiency due to loss, which results in higher photon-number being confused as having lower photon-number. In some protocols, such as type-II fusion, which relies on coincidence detection, inefficiency is not problematic. However, in most applications, including type-I fusion, this is highly problematic as photon-number confusion results in decoherence via projecting onto the wrong photon-number, unknown to the observer. Other common problems that plague these detectors are dark counts which add spurious counts from the background. APDs suffer from dead-time which is a period of time following a detection event during which it is inactive. In applications requiring feedforwarding, this is problematic. Although TESs are great detectors, they are expensive because they require cryogenic equipment to function. The multiplexed SNSPDs suffer from the same constrain and also from low quantum efficiencies. While spatially multiplexed APDs are relatively cheap, they suffer from spurious counts known as crosstalk that arises from avalanches triggered by electrons launched from neighboring APDs \cite{bib:Eraerds07,bib:Afek09,bib:Kalashnikov12}. Given the important role of detectors in linear optical quantum computation (see Section \ref{sec:LOQC}), more could be done to improve them.

@@

\subsection{Linear optics quantum computation (LOQC)}\label{sec:LOQC}
With the advancement in single-photon state preparation, detectors and quantum interferometers, quantum computation using linear optics has become feasible. 
For universal quantum computation using linear optics and single photon inputs, nonlinear couplings between the different optical modes are needed. Photons, being bosons, do not interact with one another like the way fermions do. Hence, some careful engineering has to be done in order to achieve this nonlinear coupling between modes. Some of the methods used to introduce this nonlinearity include measurements and feedforwarding \cite{bib:KLM01}, and  photon-atom interactions \cite{bib:Brod16}. These nonlinearities have been used to construct entangling gates required for (approximate) universality in quantum computation using the circuit model. They have also been used to construct entangled graph states, like the cluster states covered in Section \ref{section:cluster}, that are used in MBQC. Comprehensive reviews have been written previously to cover this topic \cite{bib:Kok05,bib:JW12}, hence we will skip some basic information to focus on recent advances. There are multiple challenges to implementing a scalable, fully-optical quantum computer, and progress has been focused on overcoming these challenges. We will discuss a few of these challenges, and the advances that have been made to overcome them.

@@

 \begin{figure}[!htb]
 \centering
\includegraphics[width=0.3\columnwidth]{phE}
\caption{A schematic diagram of the Mach-Zehnder interferometer comprising of two 50/50 beamsplitters with a phase shift $\varphi$ in one of its two paths.} \label{fig:MZ}	
\end{figure}
